This thesis marks an important milestone in my journey at the Politecnico di Milano. 
In the last four years, I have gained extensive experience, studied a wide range of topics, and conducted my research. 
This achievement would not have been possible without the technical and personal support of many people.

First of all, I would like to thank my supervisor, Prof. Livio Pinto, who has been guiding me since my BSc thesis. 
He introduced me to photogrammetry, 3D reconstruction, Macugnaga and Belvedere Glacier, which has become a second home to me. 
I am deeply grateful for his support throughout my PhD research and the writing process of this thesis.
His dedication to my academic and personal growth has been precious and will continue to shape my career for years to come.

I would also like to thank my co-supervisor, Prof. Francesco Nex, for believing in the potential of the project of a photogrammetric monitoring system with low-cost stereo cameras.
Hosting me at the University of Twente allowed me to tackle challenging problems and delve into state-of-the-art topics such as deep learning feature matching and reconstruction algorithms. 
His ongoing encouragement and patience during the final stages of paper and thesis writing have been truly invaluable.

I would also like to thank the crew of the LabMGF (Laboratory of Geodetic and Photogrammetric Measurements) research group: Federica, Federico, and Rebecca. 
Our collaboration went beyond mere work and became a true friendship. 
Thanks for the support and for the numerous projects and fieldwork, including all the surveys of bridges, dams, and, of course, Belvedere Glacier.
A special acknowledgment goes to Federica, with whom I have worked since our first year at university.
Thanks for sharing countless experiences, thoughts, complaints, and suggestions for the future and for the many conference trips.
Thanks also to Öykü and Chiara for their invaluable friendship, support for pushing towards the end of this thesis, and all the moments spent together in the lab.

I am grateful to Luca Morelli, with whom I have worked closely over the past year, for sharing our passion for developing Deep-Image-Matching.
My thanks also go to Niccolò Dematteis and Daniele Giordan for their enthusiasm and valuable contributions to the stereo monitoring project of the Belvedere Glacier, particularly in the development of our latest paper.

To all the friends I have met at various conferences over the past four years, thank you for the valuable discussions that have helped improve our research. 
I am especially grateful to Vivien, whom I met at our first EGU participation and we have developed a friendship that goes beyond professional collaboration.

Finally, I thank the reviewers, Prof. Anette Eltner and Prof. Riccardo Roncella, for their suggestions that improved my thesis.

\vspace{1pt}

\begin{flushright}
    \textit{Francesco}
\end{flushright}